---
title: "03 Variables"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating Variables in R

Variables are containers for storing data values.

R does not have a command for declaring a variable. A variable is created the moment you first assign a value to it. To assign a value to a variable, use the `<-` sign. To output (or print) the variable value, just type the variable name:

```{r}
name <- "John"
age <- 40

name   # output "John"
age    # output 40
```

From the example above, `name` and `age` are variables, while `"John"` and `40` are values.

In other programming language, it is common to use `=` as an assignment operator. In R, we can use both `=` and `<-` as assignment operators.

However, `<-` is preferred in most cases because the `=` operator can be forbidden in some context in R.

## Print / Output Variables

Compared to many other programming languages, you do not have to use a function to print/output variables in R. You can just type the name of the variable:

```{r}
name <- "John Doe"

name # auto-print the value of the name variable
```

However, R does have a `print()` function available if you want to use it. This might be useful if you are familiar with other programming languages, such as Python, which often use a `print()` function to output variables.

```{r}
name <- "John Doe"

print(name) # print the value of the name variable
```

And there are times you must use the `print()` function to output code, for example when working with `for` loops (which you will learn more about in a later chapter):

```{r}
for (x in 1:10) {
  print(x)
}
```

Conclusion: It is up to your if you want to use the `print()` function or not to output code. However, when your code is inside an R expression (for example inside curly braces `{}` like in the example above), use the `print()` function if you want to output the result.

## Concatenate Elements

You can also concatenate, or join, two or more elements, by using the `paste()` function.

To combine both text and a variable, R uses comma (`,`):

```{r}
text <- "awesome"

paste("R is", text)
```

You can also use `,` to add a variable to another variable:

```{r}
text1 <- "R is"
text2 <- "awesome"

paste(text1, text2)
paste0(text1, text2)
```

For numbers, the `+` character works as a mathematical operator:

```{r}
num1 <- 5
num2 <- 10

num1 + num2
```

If you try to combine a string (text) and a number, R will give you an error:

```{r}
num <- 5
text <- "Some text"

# The below doesn't work
# num + text
```

## Multiple Variables

R allows you to assign the same value to multiple variables in one line:

```{r}
# Assign the same value to multiple variables in one line
var1 <- var2 <- var3 <- "Orange"

# Print variable values
var1
var2
var3
```

## Variable Names

A variable can have a short name (like x and y) or a more descriptive name (age, carname, total_volume). Rules for R variables are:

-   A variable name must start with a letter and can be a combination of letters, digits, period(.) and underscore(\_). If it starts with period(.), it cannot be followed by a digit.
-   A variable name cannot start with a number or underscore (\_)
-   Variable names are case-sensitive (age, Age and AGE are three different variables)
-   Reserved words cannot be used as variables (TRUE, FALSE, NULL, if...)

```{r}
# Legal variable names:
myvar <- "John"
my_var <- "John"
myVar <- "John"
MYVAR <- "John"
myvar2 <- "John"
.myvar <- "John"

# Illegal variable names:
# 2myvar <- "John"
# my-var <- "John"
# my var <- "John"
# _my_var <- "John"
# my_v@ar <- "John"
# TRUE <- "John"
```
