---
title: "05 Numbers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Numbers

There are three number types in R:

- `numeric`
- `integer`
- `complex`

Variables of number types are created when you assign a value to them:

```{r}
x <- 10.5   # numeric
y <- 10L    # integer
z <- 1i     # complex
```

## Numeric

A `numeric` data type is the most common type in R, and contains any number with or without a decimal, like: 10.5, 55, 787:

```{r}
x <- 10.5
y <- 55

# Print values of x and y
x
y

# Print the class name of x and y
class(x)
class(y)
```

## Integer

Integers are numeric data without decimals. This is used when you are certain that you will never create a variable that should contain decimals. To create an `integer` variable, you must use the letter `L` after the integer value:

```{r}
x <- 1000L
y <- 55L

# Print values of x and y
x
y

# Print the class name of x and y
class(x)
class(y)
```

## Complex

A `complex` number is written with an "`i`" as the imaginary part:

```{r}
x <- 3+5i
y <- 5i

# Print values of x and y
x
y

# Print the class name of x and y
class(x)
class(y)
```

## Type Conversion

You can convert from one type to another with the following functions:

- `as.numeric()`
- `as.integer()`
- `as.complex()`

```{r}
x <- 1L # integer
y <- 2 # numeric

# convert from integer to numeric:
a <- as.numeric(x)

# convert from numeric to integer:
b <- as.integer(y)

# print values of x and y
x
y

# print the class name of a and b
class(a)
class(b)
```


