---
title: "06 Math"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simple Math

In R, you can use operators to perform common mathematical operations on numbers.

The `+` operator is used to add together two values:

```{r}
10 + 5
```

And the `-` operator is used for subtraction:

```{r}
10 - 5
```

## Built-in Math Functions

R also has many built-in math functions that allows you to perform mathematical tasks on numbers.

For example, the `min()` and `max()` functions can be used to find the lowest or highest number in a set:

```{r}
max(5, 10, 15)

min(5, 10, 15)
```

## sqrt()

The `sqrt()` function returns the square root of a number:

```{r}
sqrt(16)
```

## abs()

The abs() function returns the absolute (positive) value of a number:

```{r}
abs(-4.7)
```

## ceiling() and floor()

The `ceiling()` function rounds a number upwards to its nearest integer, and the `floor()` function rounds a number downwards to its nearest integer, and returns the result:

```{r}
ceiling(1.4)

floor(1.4)
```