---
title: "09 Operators"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Operators

Operators are used to perform operations on variables and values.

In the example below, we use the `+` operator to add together two values:

```{r}
10 + 5
```

R divides the operators in the following groups:

- Arithmetic operators
- Assignment operators
- Comparison operators
- Logical operators
- Miscellaneous operators

## R Arithmetic Operators

Arithmetic operators are used with numeric values to perform common mathematical operations:

```{r, echo=FALSE}
library(knitr)
data <- data.frame(Operator=c('+', '-', '*', '/', '^', '%%', '%/%'),
                   Name=c('Addition', 'Subtraction', 'Multiplication', 'Division', 'Exponent', 'Module(Remainder from division)', 'Integer Division'),
                   Example=c('x + y', 'x - y', 'x * y', 'x / y', 'x ^ y', 'x %% y', 'x%/%y'))
kable(data)
```

## R Assignment Operators

Assignment operators are used to assign values to variables:

Note: `<<-` is a global assigner. You will learn more about this in the Global Variable chapter.
It is also possible to turn the direction of the assignment operator.
x <- 3 is equal to 3 -> x

```{r}
my_var <- 3
my_var <<- 3
3 -> my_var
3 ->> my_var
my_var # print my_var
```

## R Logical Operators

Logical operators are used to combine conditional statements:

```{r, echo=FALSE}
data <- data.frame(Operator=c('&', '&&', '|', '||', '!'),
                   Description=c('Element-wise Logical AND operator. It returns TRUE if both elements are TRUE',
                                 'Logical AND operator - Returns TRUE if both statements are TRUE',
                                 'Elementwise- Logical OR operator. It returns TRUE if one of the statement is TRUE',
                                 'Logical OR operator. It returns TRUE if one of the statement is TRUE.',
                                 'Logical NOT - returns FALSE if statement is TRUE'))
kable(data)
```

## R Miscellaneous Operators

Miscellaneous operators are used to manipulate data:

```{r, echo=FALSE}
data <- data.frame(Operator=c(':', '%in%', '%*%'),
                   Description=c('Creates a series of numbers in a sequence',
                                 'Find out if an element belongs to a vector',
                                 'Matrix Multiplication'),
                   Example=c('x <- 1:10',
                             'x %in% y',
                             'x <- Matrix1 %*% Matrix2'))
kable(data)
```