---
title: "11 While Loop"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loops

Loops can execute a block of code as long as a specified condition is reached.

Loops are handy because they save time, reduce errors, and they make code more readable.

R has two loop commands:

- `while` loops
- `for` loops

## R While Loops

With the `while` loop we can execute a set of statements as long as a condition is TRUE:

```{r}
i <- 1
while (i < 6) {
  print(i)
  i <- i + 1
}
```

In the example above, the loop will continue to produce numbers ranging from 1 to 5. The loop will stop at 6 because `6 < 6` is FALSE.

The `while` loop requires relevant variables to be ready, in this example we need to define an indexing variable, `i`, which we set to 1.

## Break

With the `break` statement, we can stop the loop even if the while condition is TRUE:

```{r}
i <- 1
while (i < 6) {
  print(i)
  i <- i + 1
  if (i == 4) {
    break
  }
}
```

The loop will stop at 3 because we have chosen to finish the loop by using the `break` statement when `i` is equal to 4 (`i == 4`).

## Next

With the `next` statement, we can skip an iteration without terminating the loop:

```{r}
i <- 0
while (i < 6) {
  i <- i + 1
  if (i == 3) {
    next
  }
  print(i)
}
```

When the loop passes the value 3, it will skip it and continue to loop.

## Yahtzee!

### If .. Else Combined with a While Loop

To demonstrate a practical example, let us say we play a game of Yahtzee!

```{r}
dice <- 1
while (dice <= 6) {
  if (dice < 6) {
    print("No Yahtzee")
  } else {
    print("Yahtzee!")
  }
  dice <- dice + 1
}
```
