---
title: "12 For Loop"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## For Loops

A `for` loop is used for iterating over a sequence:

```{r}
for (x in 1:10) {
  print(x)
}
```

This is less like the `for` keyword in other programming languages, and works more like an iterator method as found in other object-orientated programming languages.

With the `for` loop we can execute a set of statements, once for each item in a vector, array, list, etc..

```{r}
fruits <- list("apple", "banana", "cherry")

for (x in fruits) {
  print(x)
}
```

```{r}
dice <- c(1, 2, 3, 4, 5, 6)

for (x in dice) {
  print(x)
}
```

The `for` loop does not require an indexing variable to set beforehand, like with `while` loops.

## Break

With the `break` statement, we can stop the loop before it has looped through all the items:

```{r}
fruits <- list("apple", "banana", "cherry")

for (x in fruits) {
  if (x == "cherry") {
    break
  }
  print(x)
}
```

The loop will stop at "cherry" because we have chosen to finish the loop by using the `break` statement when `x` is equal to "cherry" (`x == "cherry"`).

## Next

With the `next` statement, we can skip an iteration without terminating the loop:

```{r}
fruits <- list("apple", "banana", "cherry")

for (x in fruits) {
  if (x == "banana") {
    next
  }
  print(x)
}
```

When the loop passes "banana", it will skip it and continue to loop.

## Yahtzee!
### If .. Else Combined with a For Loop

To demonstrate a practical example, let us say we play a game of Yahtzee!

```{r}
dice <- 1:6

for(x in dice) {
  if (x == 6) {
    print(paste("The dice number is", x, "Yahtzee!"))
  } else {
    print(paste("The dice number is", x, "Not Yahtzee"))
  }
}
```

If the loop reaches the values ranging from 1 to 5, it prints "No Yahtzee" and its number. When it reaches the value 6, it prints "Yahtzee!" and its number.

## Nested Loops

You can also have a loop inside of a loop:

```{r}
adj <- list("red", "big", "tasty")

fruits <- list("apple", "banana", "cherry")
  for (x in adj) {
    for (y in fruits) {
      print(paste(x, y))
  }
}
```