---
title: "13 Functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating a Function

To create a function, use the `function()` keyword:

```{r}
my_function <- function() { # create a function with the name my_function
  print("Hello World!")
}
```

## Call a Function

To call a function, use the function name followed by parenthesis, like my_function():

```{r}
my_function() # call the function named my_function
```

## Arguments

Information can be passed into functions as arguments.

Arguments are specified after the function name, inside the parentheses. You can add as many arguments as you want, just separate them with a comma.

The following example has a function with one argument (fname). When the function is called, we pass along a first name, which is used inside the function to print the full name:

```{r}
my_function <- function(fname) {
  paste(fname, "Griffin")
}

my_function("Peter")
my_function("Lois")
my_function("Stewie")
```

## Number of Arguments

By default, a function must be called with the correct number of arguments. Meaning that if your function expects 2 arguments, you have to call the function with 2 arguments, not more, and not less:

```{r}
my_function <- function(fname, lname) {
  paste(fname, lname)
}

my_function("Peter", "Griffin")
```

## Default Parameter Value

The following example shows how to use a default parameter value.

If we call the function without an argument, it uses the default value:

```{r}
my_function <- function(country = "Norway") {
  paste("I am from", country)
}

my_function("Sweden")
my_function("India")
my_function() # will get the default value, which is Norway
my_function("USA")
```

## Return Values

To let a function return a result, use the `return()` function:

```{r}
my_function <- function(x) {
  return (5 * x)
}

print(my_function(3))
print(my_function(5))
print(my_function(9))
```

## Nested Functions

There are two ways to create a nested function:

- Call a function within another function.
- Write a function within a function.

```{r}
Nested_function <- function(x, y) {
  a <- x + y
  return(a)
}

Nested_function(Nested_function(2,2), Nested_function(3,3))
```

```{r}
Outer_func <- function(x) {
  Inner_func <- function(y) {
    a <- x + y
    return(a)
  }
  return (Inner_func)
}
output <- Outer_func(3) # To call the Outer_func
output(5)
```

## Recursion

R also accepts function recursion, which means a defined function can call itself.

Recursion is a common mathematical and programming concept. It means that a function calls itself. This has the benefit of meaning that you can loop through data to reach a result.

The developer should be very careful with recursion as it can be quite easy to slip into writing a function which never terminates, or one that uses excess amounts of memory or processor power. However, when written correctly, recursion can be a very efficient and mathematically-elegant approach to programming.

In this example, `tri_recursion()` is a function that we have defined to call itself ("recurse"). We use the `k` variable as the data, which decrements (`-1`) every time we recurse. The recursion ends when the condition is not greater than 0 (i.e. when it is 0).

To a new developer it can take some time to work out how exactly this works, best way to find out is by testing and modifying it.

```{r}
tri_recursion <- function(k) {
  if (k > 0) {
    result <- k + tri_recursion(k - 1)
    print(result)
  } else {
    result = 0
    return(result)
  }
}
tri_recursion(6)
```
