---
title: "15 Data Structures - Vectors"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vectors

A vector is simply a list of items that are of the same type.

To combine the list of items to a vector, use the `c()` function and separate the items by a comma.

In the example below, we create a vector variable called fruits, that combine strings:

```{r}
# Vector of strings
fruits <- c("banana", "apple", "orange")

# Print fruits
fruits
```

In this example, we create a vector that combines numerical values:

```{r}
# Vector of numerical values
numbers <- c(1, 2, 3)

# Print numbers
numbers
```

To create a vector with numerical values in a sequence, use the `:` operator:

```{r}
# Vector with numerical values in a sequence
numbers <- 1:10

numbers
```

You can also create numerical values with decimals in a sequence, but note that if the last element does not belong to the sequence, it is not used:

```{r}
# Vector with numerical decimals in a sequence
numbers1 <- 1.5:6.5
numbers1

# Vector with numerical decimals in a sequence where the last element is not used
numbers2 <- 1.5:6.3
numbers2
```

In the example below, we create a vector of logical values:

```{r}
# Vector of logical values
log_values <- c(TRUE, FALSE, TRUE, FALSE)

log_values
```

## Vector Length

To find out how many items a vector has, use the `length()` function:

```{r}
fruits <- c("banana", "apple", "orange")

length(fruits)
```

## Sort a Vector

To sort items in a vector alphabetically or numerically, use the `sort()` function:

```{r}
fruits <- c("banana", "apple", "orange", "mango", "lemon")
numbers <- c(13, 3, 5, 7, 20, 2)

sort(fruits)  # Sort a string
sort(numbers) # Sort numbers
```

## Access Vectors

You can access the vector items by referring to its index number inside brackets `[]`. The first item has index 1, the second item has index 2, and so on:

```{r}
fruits <- c("banana", "apple", "orange")

# Access the first item (banana)
fruits[1]
```

You can also access multiple elements by referring to different index positions with the `c()` function:

```{r}
fruits <- c("banana", "apple", "orange", "mango", "lemon")

# Access the first and third item (banana and orange)
fruits[c(1, 3)]
```

You can also use negative index numbers to access all items except the ones specified:

```{r}
fruits <- c("banana", "apple", "orange", "mango", "lemon")

# Access all items except for the first item
fruits[c(-1)]
```

## Change an Item

To change the value of a specific item, refer to the index number:

```{r}
fruits <- c("banana", "apple", "orange", "mango", "lemon")

# Change "banana" to "pear"
fruits[1] <- "pear"

# Print fruits
fruits
```

## Repeat Vectors

To repeat vectors, use the `rep()` function:

```{r}
repeat_each <- rep(c(1,2,3), each = 3)

repeat_each
```

```{r}
repeat_times <- rep(c(1,2,3), times = 3)

repeat_times
```

```{r}
repeat_indepent <- rep(c(1,2,3), times = c(5,2,1))

repeat_indepent
```

## Generating Sequenced Vectors

One of the examples on top, showed you how to create a vector with numerical values in a sequence with the `:` operator:

```{r}
numbers <- 1:10

numbers
```

To make bigger or smaller steps in a sequence, use the `seq()` function:

```{r}
numbers <- seq(from = 0, to = 100, by = 20)

numbers
```

Note: The `seq()` function has three parameters: `from` is where the sequence starts, `to` is where the sequence stops, and `by` is the interval of the sequence.