---
title: "16 Data Structures - Lists"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lists

A list in R can contain many different data types inside it. A list is a collection of data which is ordered and changeable.

To create a list, use the `list()` function:

```{r}
# List of strings
thislist <- list("apple", "banana", "cherry")

# Print the list
thislist
```

## Access Lists

You can access the list items by referring to its index number, inside brackets. The first item has index 1, the second item has index 2, and so on:

```{r}
thislist <- list("apple", "banana", "cherry")

thislist[1]
```

## Change Item Value

To change the value of a specific item, refer to the index number:

```{r}
thislist <- list("apple", "banana", "cherry")
thislist[1] <- "blackcurrant"

# Print the updated list
thislist
```

## List Length

To find out how many items a list has, use the `length()` function:

```{r}
thislist <- list("apple", "banana", "cherry")

length(thislist)
```

## Check if Item Exists

To find out if a specified item is present in a list, use the `%in%` operator:

```{r}
thislist <- list("apple", "banana", "cherry")

"apple" %in% thislist
```

## Add List Items

To add an item to the end of the list, use the `append()` function:

```{r}
thislist <- list("apple", "banana", "cherry")

append(thislist, "orange")
```

To add an item to the right of a specified index, add "`after=index number`" in the `append()` function:

```{r}
thislist <- list("apple", "banana", "cherry")

append(thislist, "orange", after = 2)
```

## Remove List Items

You can also remove list items. The following example creates a new, updated list without an "apple" item:

```{r}
thislist <- list("apple", "banana", "cherry")

newlist <- thislist[-1]

# Print the new list
newlist
```

## Range of Indexes

You can specify a range of indexes by specifying where to start and where to end the range, by using the `:` operator:

```{r}
thislist <- list("apple", "banana", "cherry", "orange", "kiwi", "melon", "mango")

(thislist)[2:5]
```

Note: The search will start at index 2 (included) and end at index 5 (included).

Remember that the first item has index 1.

## Loop Through a List

You can loop through the list items by using a `for` loop:

```{r}
thislist <- list("apple", "banana", "cherry")

for (x in thislist) {
  print(x)
}
```

## Join Two Lists

There are several ways to join, or concatenate, two or more lists in R.

The most common way is to use the `c()` function, which combines two elements together:

```{r}
list1 <- list("a", "b", "c")
list2 <- list(1,2,3)
list3 <- c(list1,list2)

list3
```
