---
title: "17 Data Structures - Matrices"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Matrices

A matrix is a two dimensional data set with columns and rows.

A column is a vertical representation of data, while a row is a horizontal representation of data.

A matrix can be created with the `matrix()` function. Specify the `nrow` and `ncol` parameters to get the amount of rows and columns:

```{r}
# Create a matrix
thismatrix <- matrix(c(1,2,3,4,5,6), nrow = 3, ncol = 2)

# Print the matrix
thismatrix
```

You can also create a matrix with strings:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange"), nrow = 2, ncol = 2)

thismatrix
```

## Access Matrix Items

You can access the items by using `[ ]` brackets. The first number "1" in the bracket specifies the row-position, while the second number "2" specifies the column-position:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange"), nrow = 2, ncol = 2)

thismatrix[1, 2]
```

The whole row can be accessed if you specify a comma after the number in the bracket:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange"), nrow = 2, ncol = 2)

thismatrix[2,]
```

The whole column can be accessed if you specify a comma before the number in the bracket:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange"), nrow = 2, ncol = 2)

thismatrix[,2]
```

## Access More Than One Row

More than one row can be accessed if you use the `c()` function:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange","grape", "pineapple", "pear", "melon", "fig"), nrow = 3, ncol = 3)

thismatrix
thismatrix[c(1,2),]
```

## Access More Than One Column

More than one column can be accessed if you use the `c()` function:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange","grape", "pineapple", "pear", "melon", "fig"), nrow = 3, ncol = 3)

thismatrix
thismatrix[, c(1,2)]
```

## Add Rows and Columns

Use the `cbind()` function to add additional columns in a Matrix:

Note: The cells in the new column must be of the same length as the existing matrix.

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange","grape", "pineapple", "pear", "melon", "fig"), nrow = 3, ncol = 3)
thismatrix

newmatrix <- cbind(thismatrix, c("strawberry", "blueberry", "raspberry"))

# Print the new matrix
newmatrix
```

Use the `rbind()` function to add additional rows in a Matrix:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange","grape", "pineapple", "pear", "melon", "fig"), nrow = 3, ncol = 3)
thismatrix

newmatrix <- rbind(thismatrix, c("strawberry", "blueberry", "raspberry"))

# Print the new matrix
newmatrix
```

## Remove Rows and Columns

Use the `c()` function to remove rows and columns in a Matrix:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange", "mango", "pineapple"), nrow = 3, ncol =2)
thismatrix

#Remove the first row and the first column
thismatrix <- thismatrix[-c(1), -c(1)]

thismatrix
```

## Check if an Item Exists

To find out if a specified item is present in a matrix, use the `%in%` operator:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange"), nrow = 2, ncol = 2)

"apple" %in% thismatrix
```

## Number of Rows and Columns

Use the `dim()` function to find the number of rows and columns in a Matrix:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange"), nrow = 2, ncol = 2)

dim(thismatrix)
```

## Matrix Length

Use the `length()` function to find the dimension of a Matrix:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange"), nrow = 2, ncol = 2)

length(thismatrix)
```

Total cells in the matrix is the number of rows multiplied by number of columns.

In the example above: Dimension = 2*2 = 4.

## Loop Through a Matrix

You can loop through a Matrix using a `for` loop. The loop will start at the first row, moving right:

```{r}
thismatrix <- matrix(c("apple", "banana", "cherry", "orange"), nrow = 2, ncol = 2)
thismatrix

for (rows in 1:nrow(thismatrix)) {
  for (columns in 1:ncol(thismatrix)) {
    print(thismatrix[rows, columns])
  }
}
```

## Combine two Matrices

Again, you can use the `rbind()` or `cbind()` function to combine two or more matrices together:

```{r}
# Combine matrices
Matrix1 <- matrix(c("apple", "banana", "cherry", "grape"), nrow = 2, ncol = 2)
Matrix2 <- matrix(c("orange", "mango", "pineapple", "watermelon"), nrow = 2, ncol = 2)
Matrix1
Matrix2

# Adding it as a rows
Matrix_Combined <- rbind(Matrix1, Matrix2)
Matrix_Combined

# Adding it as a columns
Matrix_Combined <- cbind(Matrix1, Matrix2)
Matrix_Combined
```

